A Stochastic Quasi-Newton Optimizer for TensorFlow

### 2018-05-07 paper/main.tex

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{../plots/nonconvex_results}
  \caption{
    SdLBFGS compared with traditional BFGS together with AdaGrad and Stochastic
    Gradient Descent on the problem $f(x) = x^6 - 2x^5 + x^3 - x^2 + 3x$, shown
    on the right. Each algorithm is initialized near the local minimum at around
    $x = 1.6$. We see that SGD, AdaGrad, and BFGS quickly converge to the local
    minimum, whereas SdLBFGS converges to the global minimum at around $x =
    -0.8$ more slowly.
  }
  \label{fig:convex-linear-system}
\end{figure}

@@

\begin{enumerate}
  \item Explaining why the basic project --- implementing a useful practical
    TensorFlow optimzier --- failed.
  \item Takeaways from the results.
  \item Future work.
\end{enumerate}

>>>>

We evaluate our implementation of SdLBFGS on a collection of test functions.
Figure \ref{fig:convex-linear-system} shows the effect of changing the memory
capacity on a relatively small fully-observable convex differentiable problem.
Figure \ref{fig:nonconvex-comparison}.

We do not report on results of using the \texttt{SQNOptimizer} to solve
standard neural network problems in the \texttt{TensorFlow} framework.
Preliminary results suggest that the present implementation scales
insufficiently well for use in large scale machine learning problems.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{../plots/nonconvex_results}
  \caption{
    SdLBFGS compared with traditional BFGS together with AdaGrad and Stochastic
    Gradient Descent on the problem $f(x) = x^6 - 2x^5 + x^3 - x^2 + 3x$, shown
    on the right. Each algorithm is initialized near the local minimum at around
    $x = 1.6$. We see that SGD, AdaGrad, and BFGS quickly converge to the local
    minimum, whereas SdLBFGS converges to the global minimum at around $x =
    -0.8$ more slowly.
  }
  \label{fig:nonconvex-comparison}
\end{figure}

@@

In this work we sought to develop an implementation of SdLBFGS which can be used
to solve large scale machine learning problems using \texttt{TensorFlow}. By
this measure, we have reported a negative result. Our preliminary results
suggest that our implementation, broadly speaking, works satisfactorily on small
problems, but our attempts at using the \texttt{SQNOptimizer} on problems of the
size of a typical neural network have been proven infeasible.

Might we say with certainty, then, that the external optimizer approach taken in
this paper is simply insufficient for solving problems that are typical for the
\texttt{TensorFlow} framework? Much analysis of our implementation remains
before we may reject with certainty our approach. The fundamental obstacles we
encountered were sensitivities to algorithm parameter choices and overall speed
in computation, both of which yield many possibilities for future work. In the
former, there may exist choices of step size $\alpha_k$ which perform
sufficiently admirably for this application, or the trade off frontier between
memory capacity and computational feasibility may yet become apparent.

In the latter, many directions exist through which the performance of our native
Python and \texttt{NumPy} implementation might improve significantly. Having
identified and dealt with potential bottlenecks around the main body of the
SdLBFGS update, the judicious introduction of performance optimizing tools could
bring the runtime cost of the implementation into relative feasibility. Such
tools exist inside the Python ecosystem, from JIT-enhancing through Numba to
static typing optimizations with Cython. In many cases significant speedups may
be found by merely modifying the underlying data structures or by adopting
lazier evaluation techniques.

The central point is that we cannot rule out the
\texttt{ExternalOptimizerInterface} approach until the exploration of these
potential improvements has sufficiently resolved the questions raised above.
Further effort is needed to resolve this partial result.

% \begin{enumerate}
%   \item Explaining why the basic project --- implementing a useful practical
%     TensorFlow optimzier --- failed.
%   \item Takeaways from the results.
%   \item Future work.
% \end{enumerate}
